Rising from the Ashes: How the Milky Way Got Its Scars

### 2024-04-29 paper/src/dis.tex
Update intro
Also moves some stuff to other places
More changes

\section{Discussion}\label{sec:discussion}
In this work, we demonstrate that a bimodal structure in the nuclear abundance plane is associated with a brief quiescent phase. We now construct a possible framework for understanding the formation of the bimodality in our simulation, discuss its relation to prior work, and then discuss predictions and connections to observations.

@@

This drop can be understood in terms of the mechanics of the galaxy formation model as described in Section~\ref{ssec:gfm}. When a star particle forms, it only takes $\sim40\,\Myr$ before the last Type II SN has gone off and iron enrichment is dominated by Type Ia SNe. As a result, during the quiescent phase there is relatively little $\alpha$-element production. On the other hand, star particles are still able to generate iron through Type Ia SNe which have a much longer period of influence. In the TNG model, the number of Type Ia SNe decays with time as $(t/\tau_8)^{-s}$, where $\tau_8=40\,\Myr$ is the lifetime of an $8\,\Msun$ star and $s=1.12$. Therefore, the burst of star formation at $t\sim2.25\,\Gyr$ is able to generously produce iron starting at $2.5\,\Gyr$, when the drop is the most steep.

\subsection{Dilution Interpretation}\label{ssec:dilute}
In some previous work, it was reported that the bimodality is a consequence of a sudden deposition of metal-poor, $\alpha$-poor gas by a satellite or cosmological filament -- i.e., a ``dilution'' \citep{2020MNRAS.491.5435B,}. The separation of the sequences follows from the rapidity of the dilution.

@@

\subsection{Direct Search for Quiescence}\label{ssec:obsqui}
Figure~\ref{fig:before_after_sfh_by_iron} indicates a very direct observational test of the mechanism proposed in this work: for disk stars at a given \FeH{}, there should be a gap of $\sim300\,\Myr$ in ages at $\sim8\,\Gyr$. With a survey of properly chosen stars, this could potentially be achieved with a modest (few hundred) sample of stars with age uncertainties of $\sim1\%$. The present observational landscape for age dating of old stars is far from this goal. To our knowledge, the best method at these ages is differential analysis of solar twins, which can provide an age uncertainty of $\sim5\%$ \citep[e.g.][]{2014ApJ...795...23B,2018MNRAS.474.2580S}. However, this only works for stars with solar metallicity, where there is not a clear separation between the high- and low-$\alpha$ sequences (though a gap in ages may still be present).

@@

\subsection{Observational Evidence of Quiescence}\label{ssec:obshiz}
Our main argument in this paper is that the Milky Way underwent a brief quiescent phase induced by a starburst at $z\sim2$. In observations, a Milky Way-analogue would be identified as a post-starburst galaxy (PSB). With abundance matching, we expect the Milky Way's total stellar mass to be $\sim2\times10^{10}\,\Msun$. A number of authors have explored PSBs and quiescent galaxies of this mass at $z\sim2$, with large advances in the post-JWST era.

First, PSBs are not uncommon. \citet{2023ApJ...953..119P} found that in massive galaxies ($M_* > 10^{10.6}\,\Msun$) the fraction of PSBs increases from $\sim2.7\%$ ($99/3655$) at $1.0 < z < 1.44$ to $\sim8\%$ ($89/1118$) at $2.16 < z < 2.5$, confirming previous work that PSBs are more common towards $z\sim2$ \citep[e.g.,][]{2016MNRAS.463..832W}. If these galaxies can be quickly rejuvenated as the system studied in this work would suggest, then the true fraction of galaxies that go through a starburst-quenching phase may be higher.

\citet{2018MNRAS.477.5072M} found that Milky Way-like bimodalities are rare in EAGLE, occuring in $\sim5\%$ of galaxies. \citet{2021MNRAS.501..236D,2022MNRAS.515.1430D} showed that merger-induced quenching in zooms can occur in the EAGLE model \citep[see also][]{2017MNRAS.465..547P}. However, the situation may be different in the lower resolution large box.

@@

There is also the perennial problem of diffusion within the hydrodynamics solver. In purely Lagrangian solvers, there is no diffusion between resolution elements, while in Eulerian codes the diffusion can be quite high. AREPO limits the numerical diffusion by allowing the mesh to move in a quasi-Lagrangian manner, and using a second-order solver \red{cite}. In FIRE-2 \citep{2018MNRAS.480..800H}, which uses the Lagrangian code GIZMO \citep{2015MNRAS.450...53H}, a subgrid turbulent metal diffusion model was used. It would be interesting to see how models with different diffusivity properties would relax or strengthen the necessity of a quiescent period to produce a bimodality.

>>>>

\section{Discussion}\label{sec:discussion}
We have demonstrated that a bimodal structure in the nuclear abundance plane is associated with a brief quiescent phase in our idealized merger simulations. We now construct a possible framework for understanding the formation of the bimodality, discuss its relation to prior work, and then discuss predictions and connections to observations.

@@

This drop can be understood in terms of the mechanics of the galaxy formation model as described in Section~\ref{ssec:gfm}. When a star particle forms, it only takes $\sim40\,\Myr$ before the last Type II SN has gone off and Fe enrichment is dominated by Type Ia SNe. As a result, during the quiescent phase there is relatively little $\alpha$-element production. On the other hand, star particles are still able to generate Fe through Type Ia SNe which have a much longer period of influence. In the TNG model, the number of Type Ia SNe decays with time as $(t/\tau_8)^{-s}$, where $\tau_8=40\,\Myr$ is the lifetime of an $8\,\Msun$ star and $s=1.12$. Therefore, the burst of star formation at $t\sim2.25\,\Gyr$ is able to generously produce Fe starting at $2.5\,\Gyr$, when the drop in \MgFe{} is the most steep.

\citet{2018MNRAS.477.5072M} found that Milky Way-like bimodalities are rare in EAGLE, occuring in $\sim5\%$ of galaxies. \citet{2021MNRAS.501..236D,2022MNRAS.515.1430D} showed that merger-induced quenching in zooms can occur in the EAGLE model \citep[see also][]{2017MNRAS.465..547P}. However, the situation may be different in the lower resolution large box. Furthermore, if the proposed starburst-quenching phase is driven by AGN feedback, then the outcome of any particular cosmological simulation with regards to the bimodality is intimately tied to its AGN model. Unfortunately, such models are highly uncertain \red{cite}.

\subsection{Dilution Interpretation}\label{ssec:dilute}
In some previous work, it was reported that the bimodality is a consequence of a sudden deposition of metal-poor, $\alpha$-poor gas by a satellite or cosmological filament -- i.e., a ``dilution'' \citep{2020MNRAS.491.5435B,2021MNRAS.503.5846R}. The separation of the sequences follows from the rapidity of the dilution. This was elaborated upon by \citet{2021MNRAS.503.5868R} who described a setup where the low-$\alpha$ disk forms inclined relative to the high-$\alpha$ disk.

@@

\subsection{Observational Evidence of Quiescence}\label{ssec:obshiz}
Our main argument in this paper is that the Milky Way underwent a brief quiescent phase induced by a starburst at $z\sim2$. In observations, a Milky Way-analogue would then be identified as a post-starburst galaxy (PSB). With abundance matching, we expect the Milky Way's total stellar mass to be $\sim2\times10^{10}\,\Msun$ at $z\sim2$ \citep{2013ApJ...771L..35V}. A number of authors have explored PSBs and quiescent galaxies at slightly higher masses at $z\sim2$, with large advances in the post-JWST era.

First, PSBs are not uncommon. \citet{2023ApJ...953..119P} found that in massive galaxies ($M_* > 10^{10.6}\,\Msun$) the fraction of PSBs increases from $\sim2.7\%$ ($99/3655$) at $1.0 < z < 1.44$ to $\sim8\%$ ($89/1118$) at $2.16 < z < 2.5$, confirming previous work that PSBs are more common towards $z\sim2$ \citep[e.g.,][]{2016MNRAS.463..832W}. If these galaxies can be quickly rejuvenated, as the system studied in this work would suggest, then the total fraction of galaxies that go through a starburst-quenching phase may be higher.

\subsection{Direct Search for Quiescence}\label{ssec:obsqui}
Figure~\ref{fig:before_after_sfh_by_iron} indicates a very direct observational test of the mechanism proposed in this work: for disk stars at a given \FeH{}, there should be a gap of $\sim300\,\Myr$ in ages at $\sim8\,\Gyr$. With a survey of properly chosen stars, this could potentially be achieved with a modestly sized (few hundred) sample of old stars with age uncertainties of $\sim1\%$. The present observational landscape for age dating of old stars is far from this goal. To our knowledge, the best method at these ages is differential analysis of solar twins, which can provide an age uncertainty of $\sim5\%$ \citep[e.g.][]{2014ApJ...795...23B,2018MNRAS.474.2580S}. However, this only works for stars with solar metallicity, where there is not a clear separation between the high- and low-$\alpha$ sequences (though a gap in ages may still be present).

@@

@@

There is also the perennial problem of diffusion within the hydrodynamics solver. In purely Lagrangian solvers, there is no diffusion between resolution elements, while in Eulerian codes the diffusion can be quite high. AREPO limits the numerical diffusion by allowing the mesh to move in a quasi-Lagrangian manner, and using a second-order solver \citep{2010MNRAS.401..791S}. In FIRE-2 \citep{2018MNRAS.480..800H}, which uses the Lagrangian code GIZMO \citep{2015MNRAS.450...53H}, a subgrid turbulent metal diffusion model was used. It would be interesting to see how models with different diffusivity properties would relax or strengthen the necessity of a quiescent period to produce a bimodality.

### 2024-04-29 paper/src/int.tex
More changes

% Why alpha-elements are important
The distribution of elemental abundances is a high dimensional space \citep[e.g., up to 26 elements for Arcturus;][]{2024A&A...684A..85V}. However, this space is highly degenerate, and so the effective number of dimensions is much smaller -- even possibly compressed to just \FeH{} and age \citep{2019ApJ...883..177N}. Two elements in particular have received particular interest - Fe and $\alpha$-elements. Type Ia and Type II supernovae are the main contributors of Fe and $\alpha$ enrichment. Fe is broadly produced in both types, and so its abundance is a proxy for the total metallicity of a star. On the other hand, $\alpha$-elements are mainly produced in Type II supernovae. The ratio of $\alpha$-elements to Fe (\alphaFe{}) is then a measure of the relative contributions of Type Ia and II SNe to the enrichment of a parcel of gas. It has therefore become common to compress the high-dimensional abundance space to the two dimensional \alphaFe{}-\FeH{} plane.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{figure1.pdf}
  \caption{\textbf{The abundance bimodality seen in the Milky Way can be reproduced in some idealized merger simulations.} In the upper panels, we show the distribution of stars in the \MgFe-\FeH plane. The left panel shows the observed distribution in the Milky Way from ASPCAP \red{cite}, while the right two panels show two idealized merger simulations. The idealized merger simulations are nearly identical, except that the bimodal simulation has a starting radius of $129\,\kpc$, while the unimodal simulation has a starting radius of $142\,\kpc$. We emphasize that the labels ``unimodal'' and ``bimodal'' are of the \textit{outcome} of the simulation, and do not reflect a particular choice in the setup. The bottom panels show the distribution of \MgFe at fixed \FeH. The colors indicate the fixed \FeH values, which are $-0.5$, $-0.375$, $-0.25$, and $0.25$. The Milky Way (left panels) exhibits a strong bimodal distribution of \MgFe at various \FeH. The idealized merger simulation marked as bimodal (center panels) also exhibits a bimodal distribution of \MgFe, though the structure is not as strongly defined. The idealized merger simulation marked as unimodal exhibits only weak structure, if any at all.}
  \label{fig:fig1}
\end{figure*}

>>>>

% Why alpha-elements are important
The distribution of elemental abundances is a high dimensional space \citep[e.g., 32 elements in][]{2024ApJ...961L..41J}. However, this space is highly degenerate, and so the effective number of dimensions is much smaller -- even possibly compressed to just \FeH{} and age \citep{2019ApJ...883..177N}. Two elements in particular have received particular interest - Fe and $\alpha$-elements. Type Ia and Type II supernovae are the main contributors of Fe and $\alpha$ enrichment. Fe is broadly produced in both types, and so its abundance is a proxy for the total metallicity of a star. On the other hand, $\alpha$-elements are mainly produced in Type II supernovae. The ratio of $\alpha$-elements to Fe (\alphaFe{}) is then a measure of the relative contributions of Type Ia and II SNe to the enrichment of a parcel of gas. It has therefore become common to compress the high-dimensional abundance space to the two dimensional \alphaFe{}-\FeH{} plane.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{figure1.pdf}
  \caption{\textbf{The abundance bimodality seen in the Milky Way can be reproduced in some idealized merger simulations.} In the upper panels, we show the distribution of stars in the \MgFe-\FeH plane. The left panel shows the observed distribution in the Milky Way from ASPCAP DR17 \citep[][J.A.~Holtzman et al., in preparation]{2016AJ....151..144G}, while the right two panels show two idealized merger simulations. The idealized merger simulations are nearly identical, except that the bimodal simulation has a starting radius of $129\,\kpc$, while the unimodal simulation has a starting radius of $142\,\kpc$. We emphasize that the labels ``unimodal'' and ``bimodal'' are of the \textit{outcome} of the simulation, and do not reflect a particular choice in the setup. The bottom panels show the distribution of \MgFe at fixed \FeH. The colors indicate the fixed \FeH values, which are $-0.5$, $-0.375$, $-0.25$, and $0.25$. The Milky Way (left panels) exhibits a strong bimodal distribution of \MgFe at various \FeH. The idealized merger simulation marked as bimodal (center panels) also exhibits a bimodal distribution of \MgFe, though the structure is not as strongly defined. The idealized merger simulation marked as unimodal exhibits only weak structure, if any at all.}
  \label{fig:fig1}
\end{figure*}

### 2024-04-29 paper/src/met.tex
Update intro
Also moves some stuff to other places
More changes

\section{Methods}\label{sec:methods}
\subsection{Isolated Setup}\label{ssec:iso_setup}
In isolation, each of the central and satellite galaxies are a compound halo setup, with a Hernquist dark matter halo and a gaseous halo with a $\beta$-profile:
\begin{equation*}
\rho = \rho_0 \left[1 + \left(\frac{r}{r_c}\right)^2\right]^{-\frac{3\beta}{2}}
\end{equation*}
The total mass within the virial radius is kept fixed, and the mass of the dark matter halo and central density of the gaseous halo are chosen to satisfy a given baryon fraction $f_b$. The dark matter halo is initialized to be in gravitational equilibrium with the total potential. The gaseous halo is in gravito-hydrostatic equilibrium, where the temperature is allowed to vary as a function of radius. The azimuthal velocity of the gaseous halo is given as a fraction of the circular velocity. There is no initial stellar disk or bulge, and the gas is initially metal-free. Thus, all star particles and metals are formed self-consistently.

We manually varied the different model parameters until we arrived at a setup that resulted in reaosnable galaxies (as determined by their stellar mass). For the central (Milky Way) galaxy, we set $M_{200}=5\times10^{11}\,\Msun$, $c_{200}=4.1$, $\beta=0.8$, $r_c=9\,\kpc$, $f_b=0.08$, and $v_{\phi}/v_{\textrm{c}}=0.2$, where $c_{200}$ is the concentration and $v_{\phi}/v_{\textrm{c}}$ is the azimuthal velocity of the gaseous halo (as a fraction of the circular velocity). For the satellite (GSE) galaxy, we set $M_{200}=2.2\times10^{11}\,\Msun$, $c_{200}=4.33$, $\beta=0.8$, $r_c=6.5\,\kpc$, $f_b=0.06$, and $v_{\phi}/v_{\textrm{c}}=0.4$.

The stellar mass build-up of our Milky Way-like and GSE-like galaxies is given in Figure~\ref{fig:mass_size}. The upper panel shows the stellar mass history. We attempt to match the expected mass of the present-day thick disk \citep[$\sim6\times10^9\,\Msun$, horizontal blue dashed line][]{2016ARA&A..54..529B} at an evolution time of $\sim3\,\Gyr$ (corresponding to $z\sim2$, vertical dashed gray line).\footnote{Of course, this neglects the significant mass contribution of the bulge, which presumably formed earlier. However, our setup does not form a strong spheroidal component. Using the trick in e.g. \citet{2022MNRAS.515.1524Z}, we take the bulge mass to be twice the counterrotating stellar mass. At $3\,\Gyr$ in the isolated Milky Way-like galaxy, the bulge mass is $\sim7\times10^{8}\,\Msun$, or $\sim13\%$ of the total mass. The Milky Way's bulge is  $\sim1.5\times10^{10}\,\Msun$, although there is strong debate about just how much of the bulge is a classical bulge which formed before the disk \citep{2016ARA&A..54..529B}. In any case, we did not attempt to match any particular property of the bulge, though one could promote bulge formation by reducing the rotation of the gas in the inner region.} We get reasonably close at $\sim5\times10^9\,\Msun$ (blue line). For GSE, we use the best-fit mass from the $N$-body simulations of \citet{2021ApJ...923...92N} -- $5\times10^8\,\Msun$ (horizontal dashed orange line). For this, we slightly overestimate (orange line).

@@

\subsection{Orbital Configuration}\label{ssec:orbit_setup}
In order to combine the galaxies, we follow \citet{2021ApJ...923...92N}, and place the satellite galaxy on a retrograde orbit. In the fiducial simulation of \citet{2021ApJ...923...92N}, the satellite is placed at the virial radius ($129\,\kpc$), with the virial velocity ($129\,\kms$), and with a circularity of $0.5$. To test minor changes to the orbit, we ran a grid of simulations with $\pm10\%$ in each the starting radius and velocity, and $\pm0.1$ in the circularity, for a total of $27$ simulations. We performed each simulation for a duration of $8\,\Gyr$.\footnote{Here, and throughout, we refer to the standard native time unit $\kpc/\left(\kms\right)$ as $\Gyr$ for convenience. There is only a $\sim3\%$ difference between the two.}

@@

\subsection{Feedback and Enrichment Model}\label{ssec:gfm}
Our feedback model is most closely related to that used in the Illustris TNG simulation suite \citep{2013MNRAS.436.3031V,2017MNRAS.465.3291W,2018MNRAS.473.4077P}. In this model, gravity and magnetohydrodynamics are solved using a \citet{1986Natur.324..446B} tree coupled to a second order finite volume fluid solver in AREPO \citep{2010MNRAS.401..791S,2016MNRAS.455.1134P}. Stellar feedback is included through a subgrid wind particle model \citep{2003MNRAS.339..289S}. AGN feedback follows a dual kinetic and thermal mode for low- and high-accretion rates \citep{2017MNRAS.465.3291W}.

In this work, we made some simplifications to this model in order to aid interpretation. First, we ignore magnetic fields. This was motivated by an initial desire to understand the CGM accretion rates in terms of idealized cooling flow solutions, but we did not revisit turning them back on. In any case, it is not clear if the magnetic fields would be realistically generated given our initial setup. Second, we use a gentler wind feedback model as described in \citet{2019MNRAS.489.4233M}. Because our setup includes both an initially steep central potential and no steady-state disk, a stronger feedback model would require a higher central gas density to achieve a reasonable SFH which introduced its own set of instabilities.

In this model, star particles synthesize elements through three different channels for which we cite the relevant yield tables: SNe Ia \citep{1997NuPhA.621..467N}, SNe II \citep{1998A&A...334..505P,2006ApJ...653.1145K}, and AGB stars \citep{2010MNRAS.403.1413K,2014MNRAS.437..195D,2014ApJ...797...44F}. Recall that star particles are simple stellar populations (i.e., a collection of stars with the same age and chemical composition). Each star particle continuously injects metals into its surroundings in the following sequence\footnote{The kinetic/thermal feedback component is handled through the wind generation, which is completely separate in this model.}:
\begin{enumerate}
    \item $t\lesssim10\,\Myr$: no metal injection as the first supernova ($M\sim100\,\Msun$) has not gone off
    \item $10\,\Myr \lesssim t \lesssim 40\,\Myr$: metal injection as $8\,\Msun<M<100\,\Msun$ stars die as Type II SNe
    \item $t\gtrsim40\,\Myr$: metal injection from Type Ia SNe and AGB stars
\end{enumerate}
There are a few things to note: (1) The exact timings are metallicity-dependent. (2) The \MgFe{} of ejected gas from Type II SNe are mass/time-dependent, with more massive stars contributing more Mg than less massive stars. (3) Type II SNe contribute the vast majority of Mg ($\sim10\times$ AGB and $\sim100\times$ Type Ia SNe). Type Ia and Type II SNe contribute approximately equal amounts of Fe ($\sim3\times$ AGB). See Figure~1 from \citet{2018MNRAS.473.4077P}. (4) The number of Type Ia SNe is greater for a younger stellar population, with a power law relationship $\propto t^{-1.12}$.

\subsection{Observed Abundances}\label{ssec:obs_abund}
Our aim in this work is to demonstrate the feasibility of a mechanism for structure formation in the abundance plane. We are only making a qualitative comparison to data. Therefore, we use the ASPCAP DR17 catalog of stellar abundances (\citet{2016AJ....151..144G}; J.A.~Holtzman et al., in preparation), which is publicly available, well-established, and widely used.

We first make some quality cuts, as well as restricting our sample to dwarfs \red{check}. We require:
\begin{itemize}[noitemsep]
    \item $\textrm{SNR} > 200$,
    \item $\textrm{VSCATTER} < 1\,\kms$,
    \item STARFLAG not set,
    \item $\varpi/\sigma_{\varpi} > 1$,
    \item $\log{g} < 3.5$,
    \item $\sigma_{\log{g}} < 0.2$,
\end{itemize}
where $\varpi$ is the parallax. We use the parallax, proper motion, and radial velocity from Gaia EDR3 \citep{2016AA...595A...1G, 2021AA...649A...1G, 2021AA...649A...2L, 2021AA...653A.160S}.

We next make a selection on the angular momentum of stars in order to make a solar neighborhood selection. We assume the solar radius and azimuthal velocity are $R_0=8\,\kpc$ and $V_0=220\,\kms$ \red{cite}, and select stars which have $L_z$ within $10\%$ of the solar angular momentum. We further require that $\left|z\right| < 3\,\kpc$.

@@

\subsection{Solar Neighborhood in Simulations}\label{ssec:solarneigh}
When comparing galaxy simulations to the observed solar neighborhood, some ambiguity arises in how to make a ``solar neighborhood-like'' selection of star particles. Naturally, this selection is dependent on the posed question. In this work, this is the formation of the abundance bimodality. The Sun is known to sit near the end of the thick disk, where the thick and thin disk have comparable surface densities (the ratio of thick-to-thin is $\sim12\%$ \red{cite}). As a result, the abundance bimodality appears most strongly near the Sun -- further inwards the high-$\alpha$ sequence is more dominant and further outwards the high-$\alpha$ sequence vanishes \red{cite}.

@@

This choice of angular momentum may seem unrealistic. However, it corresponds to roughly selecting star particles with radii between $2$ and $5\,\kpc$. As discussed in Section~\ref{ssec:iso_setup}, our Milky Way-like galaxy is more compact than the Milky Way's thick disk.

>>>>

\section{Methods}\label{sec:methods}
\subsection{Isolated Setup}\label{ssec:iso_setup}
We use a modified version of the \texttt{MakeNewDisk} variant described in \citet{2023MNRAS.tmp.2070B}. In isolation, each of the central and satellite galaxies are a compound halo setup, with a \citet{1990ApJ...356..359H} dark matter halo and a gaseous halo with a $\beta$-profile:
\begin{equation*}
\rho = \rho_0 \left[1 + \left(\frac{r}{r_c}\right)^2\right]^{-\frac{3\beta}{2}}
\end{equation*}
The total mass within the virial radius is kept fixed, and the mass of the dark matter halo and central density of the gaseous halo are chosen to satisfy a given baryon fraction $f_b$. The dark matter halo is initialized to be in gravitational equilibrium with the total potential. The gaseous halo is in gravito-hydrostatic equilibrium, where the temperature is allowed to vary as a function of radius. The azimuthal velocity of the gaseous halo is given as a fraction of the circular velocity. There is no initial stellar disk or bulge, and the gas is initially metal-free. Thus, all star particles and metals are formed self-consistently.

We used the fiducial halos in \citet{2021ApJ...923...92N} as a starting point for each galaxy. We then manually varied the different model parameters until we arrived at a setup that resulted in reasonable galaxies as determined by their stellar mass. For the central (Milky Way) galaxy, we set $M_{200}=5\times10^{11}\,\Msun$, $c_{200}=4.1$, $\beta=0.8$, $r_c=9\,\kpc$, $f_b=0.08$, and $v_{\phi}/v_{\textrm{c}}=0.2$, where $c_{200}$ is the concentration and $v_{\phi}/v_{\textrm{c}}$ is the azimuthal velocity of the gaseous halo as a fraction of the local circular velocity. For the satellite (GSE) galaxy, we set $M_{200}=2.2\times10^{11}\,\Msun$, $c_{200}=4.33$, $\beta=0.8$, $r_c=6.5\,\kpc$, $f_b=0.06$, and $v_{\phi}/v_{\textrm{c}}=0.4$.

The stellar mass build-up of our Milky Way-like and GSE-like galaxies is given in Figure~\ref{fig:mass_size}. The upper panel shows the stellar mass history. We attempt to match the expected mass of the present-day thick disk \citep[$\sim6\times10^9\,\Msun$, horizontal blue dashed line][]{2016ARA&A..54..529B} at an evolution time of $\sim3\,\Gyr$ (corresponding to $z\sim2$, vertical dashed gray line).\footnote{Of course, this neglects the significant mass contribution of the bulge, which presumably formed earlier. However, our setup does not form a strong spheroidal component. Using the trick in e.g. \citet{2022MNRAS.515.1524Z}, we take the bulge mass to be twice the counterrotating stellar mass. At $3\,\Gyr$ in the isolated Milky Way-like galaxy, the bulge mass is $\sim7\times10^{8}\,\Msun$, or $\sim13\%$ of the total mass. The Milky Way's bulge is  $\sim1.5\times10^{10}\,\Msun$, although there is strong debate about just how much of the bulge is a classical bulge which formed before the disk \citep{2016ARA&A..54..529B}. In any case, we did not attempt to match any particular property of the bulge, though one could promote bulge formation by reducing the rotation of the gas in the inner region.} We get reasonably close at $\sim5\times10^9\,\Msun$ (blue line). For GSE, we use the best-fit mass from the $N$-body simulations of \citet{2021ApJ...923...92N} -- $5\times10^8\,\Msun$ (horizontal dashed orange line). For this, we slightly overestimate at $\sim6\times10^8\,\Msun$ (orange line).

@@

\subsection{Orbital Configuration}\label{ssec:orbit_setup}
In order to combine the galaxies, we follow \citet{2021ApJ...923...92N}, and place the satellite on a retrograde orbit. In the fiducial simulation of \citet{2021ApJ...923...92N}, the satellite is placed at the virial radius ($129\,\kpc$), with the virial velocity ($129\,\kms$), and with a circularity of $0.5$. To test minor changes to the orbit, we ran a grid of simulations with $\pm10\%$ in each the starting radius and velocity, and $\pm0.1$ in the circularity, for a total of $27$ simulations. We performed each simulation for a duration of $8\,\Gyr$, and used \texttt{FOF} and \texttt{SUBFIND} in order to identify substructure \citep{2005Natur.435..629S,2009MNRAS.399..497D}.

@@

We show the bimodal and unimodal orbits in Figure~\ref{fig:orbit}. The centers of  The upper and middle panels show the orbits in the $x$-$y$ and $x$-$z$ planes, respectively. The lower panel shows the separation distance as a function of time. The orbits are initially retrograde, but quickly radialize after the first pericentric passage. Coalescence occurs rapidly after the second pericentric passage at $\sim2\,\Gyr$. Soon after, \texttt{SUBFIND} does not recognize the satellite as a separate subhalo as the two galaxies coalesce.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{orbit.pdf}
    \caption{The orbits of the bimodal (blue) and unimodal (orange) simulations. The upper and middle panels show the orbit in the $x$-$y$ and $x$-$z$ planes, respectively. The bottom panel shows the separation distance as a function of time. The orbit begins retrograde but then radializes after the first pericentric passage. The satellite then coalesces quickly after the second pericentric passage, after about $\sim2\,\Gyr$ of evolution.}
    \label{fig:orbit}
\end{figure}

\subsection{Feedback and Enrichment Model}\label{ssec:gfm}
Our feedback model is most closely related to that used in the Illustris TNG simulation suite \citep{2013MNRAS.436.3031V,2017MNRAS.465.3291W,2018MNRAS.473.4077P}. In this model, gravity and magnetohydrodynamics are solved using a \citet{1986Natur.324..446B} tree coupled to a second order finite volume fluid solver in AREPO \citep{2010MNRAS.401..791S,2016MNRAS.455.1134P}. Stellar feedback is included through a subgrid wind particle model \citep{2003MNRAS.339..289S}. AGN feedback follows a dual kinetic and thermal mode for low- and high-accretion rates \citep{2017MNRAS.465.3291W}, though in our setup the AGN is only ever in the high-accretion mode. The central galaxy is seeded with a black hole with the typical seed mass ($8\times10^5\,\Msun$).

In this work, we made some simplifications to this model in order to aid interpretation. First, we ignore magnetic fields. This was motivated by an initial desire to understand the CGM accretion rates in terms of idealized cooling flow solutions, but we did not revisit turning them back on. In any case, it is not clear if the magnetic fields would be realistically generated given our initial setup. Second, we use a gentler wind feedback model as described in \citet{2019MNRAS.489.4233M}. Because our setup includes both an initially steep central potential and no steady-state disk, a stronger feedback model would require a higher central gas density to achieve a reasonable SFH which introduced its own set of pathological instabilities.

In this model, star particles synthesize elements through three different channels for which we cite the relevant yield tables: SNe Ia \citep{1997NuPhA.621..467N}, SNe II \citep{1998A&A...334..505P,2006ApJ...653.1145K}, and AGB stars \citep{2010MNRAS.403.1413K,2014MNRAS.437..195D,2014ApJ...797...44F}. Recall that star particles are simple stellar populations (i.e., a collection of stars with the same age and chemical composition). Each star particle continuously injects metals into its surroundings in the following sequence\footnote{The kinetic/thermal feedback component is handled through the wind generation, which is completely separate in this model.}:
\begin{enumerate}
    \item $t\lesssim10\,\Myr$: no metal injection as the first supernova ($M\sim100\,\Msun$) has not gone off
    \item $10\,\Myr \lesssim t \lesssim 40\,\Myr$: metal injection as $8\,\Msun<M<100\,\Msun$ stars die as Type II SNe
    \item $t\gtrsim40\,\Myr$: metal injection from Type Ia SNe and AGB stars
\end{enumerate}
There are a few things to note: (1) The exact timings are metallicity-dependent. (2) The \MgFe{} of ejected gas from Type II SNe is mass/time-dependent, with more massive stars contributing more Mg than less massive stars. (3) In a Hubble time, type II SNe contribute the vast majority of Mg ($\sim10\times$ AGB and $\sim100\times$ Type Ia SNe). Type Ia and Type II SNe contribute approximately equal amounts of Fe (each $\sim3\times$ AGB). See Figure~1 from \citet{2018MNRAS.473.4077P}. (4) The number of Type Ia SNe is greater for a younger stellar population, with a power law relationship $\propto \left(t/\tau_8\right)^{-1.12}$, where $\tau_8=40\,\Myr$ is the lifetime of an $8\,\Msun$ star.

\subsection{Observed Abundances}\label{ssec:obs_abund}
Our aim in this work is to demonstrate the feasibility of a mechanism for structure formation in the abundance plane. We are only making a qualitative comparison to data. Therefore, we use the ASPCAP DR17 catalog of stellar abundances \citep[][J.A.~Holtzman et al., in preparation]{2016AJ....151..144G}, which is publicly available, well-established, and widely used.

We first make some quality cuts, as well as restricting our sample to giants. We require:
\begin{itemize}[noitemsep]
    \item $\textrm{SNR} > 200$,
    \item $\textrm{VSCATTER} < 1\,\kms$,
    \item STARFLAG not set,
    \item $\varpi/\sigma_{\varpi} > 1$,
    \item $\log{g} < 3.5$,
    \item $\sigma_{\log{g}} < 0.2$,
\end{itemize}
where $\varpi$ is the parallax. We use the parallax, proper motion, and radial velocity from Gaia EDR3 \citep{2016AA...595A...1G,2021AA...649A...1G,2021AA...649A...2L,2021AA...653A.160S}.

We next make a selection on the angular momentum of stars in order to make a solar neighborhood selection. We assume the solar radius and azimuthal velocity are $R_0=8\,\kpc$ and $V_0=220\,\kms$ \citep{2016ARA&A..54..529B}, and select stars which have $L_z$ within $10\%$ of the solar angular momentum. We further require that $\left|z\right| < 3\,\kpc$.

@@

\subsection{Solar Neighborhood in Simulations}\label{ssec:solarneigh}
When comparing galaxy simulations to the observed solar neighborhood, some ambiguity arises in how to make a ``solar neighborhood-like'' selection of star particles. Naturally, this selection is dependent on the posed question. In this work, this is the formation of the abundance bimodality. The Sun is known to sit near the end of the thick disk, where the thick and thin disk have comparable surface densities \citep[the ratio of thick-to-thin is $\sim12\%$][]{2016ARA&A..54..529B}. As a result, the abundance bimodality appears most strongly near the Sun -- further inwards the high-$\alpha$ sequence is more dominant and further outwards the high-$\alpha$ sequence vanishes \citep[e.g.,][]{2015ApJ...808..132H}.

@@

This choice of angular momentum may seem unrealistic. However, it corresponds to roughly selecting star particles with radii between $2$ and $5\,\kpc$. As discussed in Section~\ref{ssec:iso_setup}, our Milky Way-like galaxy at $3\,\Gyr$ is more compact than the Milky Way's thick disk.
